Making Q-learning robust to time discretization

### 2019-01-17 paper/icml_drau/advup.tex
.

Contrary to this discrete action exploration scheme, $\varepsilon$-greedy
exploration is likely to not explore, i.e. to collapse to a deterministic
policy when $\deltat$ goes to $0$. Indeed, one can prove that in a near
continuous deterministic environment, the exploration policy resulting from
applying $\varepsilon$ greedy exploration to a constant exploitation policy
converges to a determinist policy as $\deltat$ goes to $0$. More precisely, let
$(a_\deltat)$ a sequence of i.i.d. actions of distribution $p_A$, (for instance
the sequence of actions generated by $\varepsilon$-greedy exploration on a
constant policy). Then, as $\deltat$ goes to $0$, trajectories generated by
interacting with the environment using the $a$'s converge to solutions of
\begin{equation}
	ds_t = \E_{a\sim p_A}\left[F(s_t, a)\right]dt
\end{equation}
where the expectation cancels the exploration noise.

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous~\cite{ddpg} and
discrete~\cite{dqn} action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$,
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}
\end{align}
for a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s,
a')$. The argmax is exactly implemented in the case of discrete actions, and approximated by
a neural network $\pi_\phi(s)$, trained to maximize
$A(s, \pi_\phi(s))$ in the continuous case.  This is similar to what is done in \cite{ddpg}.

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$ of
an exploratory trajectory are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section. \TODO{Change notations for the learning rates in the algorithm}
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
In practice, $A$ and $V$ are learned using
transitions sampled from a replay buffer instead of trajectory samples, to
prevent excessive correlation between training samples~\cite{dqn}.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$:
\begin{theorem}
  If $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A\deltat$, the continuous limit trajectory for parameters follows an SDE.
\end{theorem}

>>>>

Contrary to this discrete action exploration scheme, $\varepsilon$-greedy
exploration is likely to not explore, i.e. to collapse to a deterministic
policy when $\deltat$ goes to $0$. Indeed, one can prove that in a near
continuous deterministic environment, the exploration policy resulting from
applying $\varepsilon$ greedy exploration to a constant exploitation policy
converges to a determinist policy as $\deltat$ goes to $0$. More precisely, let
$(a_\deltat)$ a sequence of i.i.d. actions of distribution $p_A$, (for instance
the sequence of actions generated by $\varepsilon$-greedy exploration on a
constant policy). Then, as $\deltat$ goes to $0$, trajectories generated by
interacting with the environment using the $a$'s converge to solutions of
\begin{equation}
	ds_t = \E_{a\sim p_A}\left[F(s_t, a)\right]dt
\end{equation}
where the expectation cancels the exploration noise. The proof is given in the
supplementary material.

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous and
discrete action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$, with a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s, a')$
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}.
\end{align}
Maximization over actions is implemented exactly in the discrete actions case,
and approximated by a policy neural network $\pi_\phi(s)$, trained to maximize $A(s,
\pi_\phi(s))$ in the continuous case.  This is similar to what is done in
\cite{ddpg}.

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$, either from
an exploratory trajectory or from a replay buffer~\cite{dqn} are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section. \TODO{Change notations for the learning rates in the algorithm}
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$.
	\end{itemize}
\end{theorem}
